Turbulence and Particle Acceleration in Giant Radio Halos: the Origin of Seed Electrons

### 2016-04-08 paper.tex
Added files via upload

\renewcommand{\vec}{\ensuremath{\mathbfit}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\Vph}{V_\mathrm{ph}}
\newcommand{\mug}{\mu G}
\newcommand{\RH}{R_\rmn{RH}}
\newcommand{\bvel}{\ensuremath{\boldsymbol{\varv}}}
\newcommand{\bnabla}{\ensuremath{\boldsymbol{\nabla}}}

@@

In our Galaxy, the CRe-to-CRp ratio at a few GeV is $K_{\rmn{ep}} \sim
10^{-2}$. Hence, we adopt this as a fiducial value for the CRe-to-CRp
acceleration efficiency (see \cite{pinzke13} for more
discussion). However, as recent PIC simulations have shown, this is
likely very different at weak shocks, with electrons efficiently
accelerated at perpendicular shocks
\citep{2014ApJ...794..153G,2014ApJ...797...47G}, and ions efficiently
accelerated at parallel shocks \citep{2014ApJ...783...91C}. Thus,
depending on magnetic geometry, $K_{\rmn{ep}}$ could be either larger
or smaller. In this work we use a simple test-particle model for the
CRp acceleration, where we adopt a fit to Monte Carlo simulations of
the thermal leakage process that relates the momentum of injected
protons to the thermal energy of the shocked plasma \citep{kang11}. We
set an upper limit on the ratio of accelerated proton-to-dissipated
energy in the downstream of strong shocks that varies from 1-10\%,
depending on the adopted model (for more details, see Section
\ref{sec:results}). In order not to violate energy conservation with a
fixed limit for the acceleration efficiency, we steepen the spectral
index of the accelerated CRps at the limit \citep{pinzke13}, which
mimics the non-linear CR modification of the shock
\citep{ellison00}. For weak shocks the efficiency is a factor 10-100
lower and well below the limit. Some observations of radio relics
suggest high values of $K_{\rmn{ep}}$, due to the absence of gamma-ray
emission, which probes the CRp population
\citep{2014MNRAS.437.2291V}. This suggests primary CRes as a viable
alternative scenario to secondary CRes as seeds for the giant RHs. In
our {\em M-primaries} scenario, the injected distribution of CRes is
derived in the same way as for the CRps. Once they have been
accelerated to relativistic energies, injected electrons and protons
are indistinguishable. We therefore assume that CRp and CRe have the
same distribution function $f_\rmn{CRe}(p) = K_\rmn{ep}
f_\rmn{CRp}(p)$, with a different normalization (due to differing
acceleration efficiencies) $K_\rmn{ep}=0.1$ \citep[which is viable for
  primarily perpendicular shocks][]{2014ApJ...794..153G}. While this
appears to contradict the radial bias of magnetic fields in the bulk
of the ICM as suggested by observations \citep{2010NatPh...6..520P}
and cluster simulations \citep{2011ApJ...740...81R}, it does not
necessarily apply to the accretion shock regions, which show a field
geometry with a net perpendicular bias with respect to the shock
normal---at least for giant radio relics \citep{2010Sci...330..347V}.

>>>>

\renewcommand{\vec}{\ensuremath{\mathbfit}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\Vph}{V_\mathrm{ph}}
\newcommand{\mug}{\mu G}
\newcommand{\RH}{R_\rmn{RH}}
\newcommand{\bvel}{\ensuremath{\boldsymbol{\varv}}}
\newcommand{\bnabla}{\ensuremath{\boldsymbol{\nabla}}}
\newcommand\eb{\epsilon_\rmn{B}}

@@

\AP{We use a simple test-particle model for the CR acceleration and
  injection, where each shock inject CRs that trace a power-law in
  momentum,
  \begin{equation}
    f(p,t) = C(t)\,p^{\alpha_\rmn{inj}}\,,
  \end{equation}
  determined by the normalization $C(t)$ and the spectral index
  $\alpha_\rmn{inj}=\frac{(\gamma_\rmn{ad}+1)\mathcal{M}^2}{(\gamma_\rmn{ad}-1)\mathcal{M}^2+2}$
  that depends on the adiabatic index $\gamma_\rmn{ad} = 5/3$ and the
  Mach number of the shock $\mathcal{M}$. It is given by the ratio of
  the upstream velocity ($u_2$) and the sound speed ($c_s$). The CR
  number density and CR energy density are derived from $n_\rmn{CR} =
  \int_{p_\rmn{inj}}^\infty \dd p\, f(p)$ and $\epsilon_\rmn{CR} =
  \int_{p_\rmn{inj}}^\infty \dd p\, f(p) \,T(p)$, respectively, where
  $T(p) = (\sqrt{1+p^2} -1)\, m\,c^2$ is the kinetic energy of a
  particle with momentum $p$. We adopt a fit to Monte Carlo
  simulations of the thermal leakage process that relates the momentum
  of injected protons ($p_\rmn{inj}$) to the thermal energy
  ($p_\rmn{th}$) of the shocked plasma \citep{kang11}:
  \begin{eqnarray}
  \label{eq:qinj}
  p_\rmn{inj} &=& x_\rmn{inj} p_\rmn{th} = x_\rmn{inj} \sqrt{\frac{2 \,k T_2}{m c^2}}\,, \nonumber \\
  \rmn{where}\quad x_\rmn{inj} &\approx& 1.17 \frac{u_2}{p_{\rmn th}\,c} \left(1+
  \frac{1.07}{\epsilon_B}\right) \left(\frac{\mathcal{M}}{3}\right)^{0.1}\,.
  \end{eqnarray}
Here $\eb = B_0/B_{\perp}$, $B_0$ is the amplitude of the downstream
MHD wave turbulence, and $B_{\perp}$ is the magnetic field along the
shock normal. The physical range of $\eb$ is quite uncertain due to
complex plasma interactions. In this paper, we adopt $\eb = 0.23$,
which we will later see corresponds to a conservative maximum energy
acceleration efficiency for protons of $10\%$. To derive the
acceleration efficiency, $\zeta_{\rmn inj}$, we first have to infer the
particle injection efficiency, which is the fraction of downstream
thermal gas particles which experience diffusive shock acceleration
\citep[for details see ][]{pinzke13},
\begin{equation}
  \label{eq:eta}
  \eta_\rmn{CR,lin} =
  \frac{4}{\sqrt{\upi}}\,\frac{x_\rmn{inj}^3}{\alpha_\rmn{inj}-1}\,
  \rmn{e}^{-x_\rmn{inj}^2}.
\end{equation}
The particle injection efficiency is a strong function of
$x_\rmn{inj}$ that depends on both $\mathcal{M}$ and $\eb$. The
energy density of CRs that are injected and accelerated at the shock
(neglecting the CR back reaction on the shock) is given by
\begin{equation}
\label{eq:CR_energy} 
  \Delta\epsilon_\rmn{CR,lin} =
  \eta_\rmn{CR,lin}(\mathcal{M})\,T_\rmn{CR}(\mathcal{M},p_\rmn{inj})\,n_\rmn{th}(T_2),
\end{equation}
and the CR energy injection and acceleration efficiency is:
\begin{equation}
  \zeta_\rmn{lin} =
  \frac{\Delta\epsilon_\rmn{CR,lin}}{\Delta\epsilon_\rmn{diss}},
   \quad\mbox{where}\quad
  \Delta\epsilon_\rmn{diss} = \epsilon_\rmn{th2} - \epsilon_\rmn{th0}\,\left(\frac{\rho_2}{\rho_0}\right)^{\gamma_\rmn{ad}}\,.
\label{eqn:energy_frac}  
\end{equation}
The dissipated energy density in the downstream regime,
$\Delta\epsilon_\rmn{diss}$, is given by the difference of the thermal
energy densities in the pre- and post-shock regimes, corrected for the
adiabatic energy increase due to gas compression.}

\AP{We limit the acceleration efficiency to $\zeta_{\rmn max}$ by
  steepening the spectral index of the injected population
  $\alpha_\rmn{inj}$ to $\alpha_\rmn{sub}$ so that $\zeta_{\rmn lin}
  \leq \zeta_{\rmn max}$ is always fulfilled. The slope
  $\alpha_\rmn{inj}$ impact $\zeta_{\rmn inj}$ via the mean energy per
  particle given by $T_\rmn{CR} = \epsilon_\rmn{CR}/n_\rmn{CR}$. This
  procedure conserves energy and is motivated by models of non-linear
  shock acceleration where a sub-shock with a lower compression ratio
  (and hence steeper spectral index) forms
  \citep[e.g.,][]{2000ApJ...540..292E}. Given our assumed $\eb=0.23$,
  we find that for strong shocks where $\alpha \lesssim 2.3$ the
  spectral slope is steepened by a maximum of $\sim 10$ per cent in
  low temperature regimes ($kT\sim 0.1$~keV), while the steepening is
  much smaller for high temperature regimes ($kT\sim 10$~keV) that are
  more relevant for clusters. Since $p_{\rm inj}$ remains fixed, so
  does the CR number density $n_\rmn{CR}$. Hence we can solve for the
  renormalized normalization constant $C_\rmn{sub}$ using $n_\rmn{CR}$
  and Eqn.~\ref{eq:eta}:
\begin{equation}
  \label{eq:Cp_sub}
  C_\rmn{sub}=\eta_\rmn{CR,lin}\,(\alpha_\rmn{sub}-1)\,p_\rmn{inj}^{\alpha_\rmn{sub}-1}\,,
\end{equation}
where the new distribution function is given by $f(p,t)= C_\rmn{sub}
p^{-\alpha_\rmn{sub}}$.} We set an upper limit on the ratio of
accelerated proton-to-dissipated energy in the downstream of strong
shocks that varies from $\zeta_{\rmn max} \sim 1-10\%$, depending on
the adopted model (for more details, see Section \ref{sec:results}).

In our Galaxy, the CRe-to-CRp ratio at a few GeV is $K_{\rmn{ep}} \sim
10^{-2}$. Hence, we adopt this as a fiducial value for the CRe-to-CRp
acceleration efficiency (see \cite{pinzke13} for more
discussion). However, as recent PIC simulations have shown, this is
likely very different at weak shocks, with electrons efficiently
accelerated at perpendicular shocks
\citep{2014ApJ...794..153G,2014ApJ...797...47G}, and ions efficiently
accelerated at parallel shocks \citep{2014ApJ...783...91C}. Thus,
depending on magnetic geometry, $K_{\rmn{ep}}$ could be either larger
or smaller. \sout{In order not to violate energy conservation with a
  fixed limit for the acceleration efficiency, we steepen the spectral
  index of the accelerated CRps at the limit %\citep{pinzke13}
, which  mimics the non-linear CR modification of the shock
  %\citep{ellison00}
  . For weak shocks the efficiency is a factor 10-100
  lower and well below the limit.} Some observations of radio relics
suggest high values of $K_{\rmn{ep}}$, due to the absence of gamma-ray
emission, which probes the CRp population
\citep{2014MNRAS.437.2291V}. This suggests primary CRes as a viable
alternative scenario to secondary CRes as seeds for the giant RHs. In
our {\em M-primaries} scenario, the injected distribution of CRes is
derived in the same way as for the CRps. Once they have been
accelerated to relativistic energies, injected electrons and protons
are indistinguishable. We therefore assume that CRp and CRe have the
same distribution function $f_\rmn{CRe}(p) = K_\rmn{ep}
f_\rmn{CRp}(p)$, with a different normalization (due to differing
acceleration efficiencies) $K_\rmn{ep}=0.1$ \citep[which is viable for
  primarily perpendicular shocks][]{2014ApJ...794..153G}. While this
appears to contradict the radial bias of magnetic fields in the bulk
of the ICM as suggested by observations \citep{2010NatPh...6..520P}
and cluster simulations \citep{2011ApJ...740...81R}, it does not
necessarily apply to the accretion shock regions, which show a field
geometry with a net perpendicular bias with respect to the shock
normal---at least for giant radio relics \citep{2010Sci...330..347V}.
